---
title: "Text Compiler Assignment"
author: "Niveda"
date: "09-30-2026"
output: html_document
---

#Jour 689: This code, largely written by the famous **Sean Mussenden**, takes separate text files and compiles them into a single dataframe for analysis.

1)  Create a folder Week5_Text_Compiler
2)  Copy your spreadsheet index and the folder of text files into Week5_Text_Compiler
3)  Create an .Rproj file for Week5_Text_Compiler

```{r}
library(tidyverse)
library(janitor)
#install.packages("striprtf")
library(striprtf)
#install.packages("tidytext")
library(tidytext)
library(rio)
```

# Reformat .RTF files

```{r}
# Load required packages

# Set the paths for your folders
input_folder <- "./long_covid_rtf/"  # Replace with your input folder path
output_folder <- "./long_covid_raw_text/" # Replace with your output folder path

# Create output folder if it doesn't exist
if (!dir.exists(output_folder)) {
  dir.create(output_folder)
}

# Get a list of all .rtf files in the input folder
rtf_files <- list.files(path = input_folder, pattern = "\\.RTF$", full.names = TRUE)

# Convert each .rtf file to .txt
for (file in rtf_files) {
  # Extract the file name without extension
  file_name <- tools::file_path_sans_ext(basename(file))
  
  # Read the RTF content
  rtf_content <- read_rtf(file)
  
  # Create output file path
  output_file <- file.path(output_folder, paste0(file_name, ".txt"))
  
  # Write the content to a .txt file
  writeLines(rtf_content, output_file)
  
  # Print progress
  cat("Converted:", file, "to", output_file, "\n")
}

cat("Conversion complete!\n")
```

# Raw text compiler

```{r include=FALSE}
#This creates an index with the file path to the stories. And then it compiles the stories into a dataframe
#####################
# Begin SM Code #####
#####################

###
# List out text files that match pattern .txt, create DF
###

#Adjust thisline for your file name
files <- list.files("./long_covid_raw_text", pattern="*.txt") %>% 
  as.data.frame() |> 
  rename(filename = 1) |> 
  #create an matching file name
  mutate(index = str_replace_all(filename, ".txt", "")) %>%
  mutate(index = tolower(gsub("[[:punct:][:space:]|]", "", index))) |> 
  mutate(index = tolower(index)) |> 
  mutate(index = str_sub(index, 1, 25)) |> 
  distinct(index, .keep_all = TRUE)


#Join the file list to the index

final_data <- rio::import("long_covid_results_list.XLSX") |> 
  clean_names() |> 
   #create an matching file name
  mutate(index = tolower(gsub("[[:punct:][:space:]|]", "", title))) |> 
  mutate(index = tolower(index)) |> 
  mutate(index = str_sub(index, 1, 25)) |> 
  distinct(index, .keep_all = TRUE)
```

### Check for duplicate entries

```{r}

final_data |> 
  count(title) |> 
  arrange(desc(n))
```


```{r}
dupe_data <- rio::import("long_covid_results_list.XLSX") |> 
  clean_names() |> 
   #create an matching file name
  mutate(index = tolower(gsub("[[:punct:][:space:]|]", "", title))) |> 
  mutate(index = tolower(index)) |> 
  mutate(index = str_sub(index, 1, 25))
```
#Counting duplicates
```{r}
dupe_data |> 
  count(title) |> 
  arrange(desc(n)) |> 
  filter(n >1)

```
#Total duplicates
```{r}
dupe_data |> 
  count(title) |> 
  arrange(desc(n)) |> 
  filter(n >1) |> 
  summarize(sum(n))
```


#Create final index
```{r}

final_index <- final_data |> 
  inner_join(files, c("index")) |> 
#you need the actual hard-coded path on this line below to the text
  mutate(filepath = paste0("./long_covid_raw_text/", filename))

head(final_index)
```

#Fact Check

```{r}

anti_final_index <- final_data |> 
  anti_join(files, c("index"))

```

#Checking for duplicates

```{r}
final_index |> 
  count(title) |> 
  arrange(desc(n))

```

#Text compiler

```{r}
###
# Define function to loop through each text file 
###

create_article_text <- function(row_value) {
  
  #row_value is the single argument that is passed to the function
  # Take each row of the dataframe
  temp <- final_index %>%
    slice(row_value)
  
  # Store the filename for  use in constructing articles dataframe
  temp_filename <- temp$filename
  
  # Create a dataframe by reading in lines of a given textfile
  # Add a filename column 
  articles_df_temp <- read_lines(temp$filepath) %>%
    as_tibble() %>%
    mutate(filename = temp_filename)
  
  # Bind results to master articles_df
  # <<- returns to global environment
  articles_df <<- articles_df %>%
    bind_rows(articles_df_temp)
}

###
# Create elements needed to run function
###

# Create empty tibble to store results
articles_df <- tibble()
#running once to test
#create_article_text(2) 
# Create an array of numbers to loop through, from 1 to the number of rows in our index dataframe 
row_values <- 1:nrow(final_index)

###
# Execute function using lapply
# This loops through each row of the dataframe and append results to master file
###

lapply(row_values, create_article_text)

###
# Clean up articles_df and join to index dataframe
###

articles_df <- articles_df %>%
  select(filename, sentence=value) %>%
  inner_join(final_index)


# write.csv(articles_df, "..articles_df.csv")

```
```{r}
view(articles_df)
```

#Creating bigrams

#checking names in long covid articles data frame
```{r}
names(articles_df)
```

#tokenize the sentence column
```{r}
long_covid_words <- articles_df %>%
  unnest_tokens(word, sentence)
```

#Remove stopwords and count the words
```{r}
word_counts <- long_covid_words |> 
  anti_join(stop_words) |> 
  count(word, sort = TRUE)

word_counts
# Word Count
#Top 10 words: covid is the word with the highest count (n=1122), followed by health (n=952), 90 (n=730), 78 (n=573), people (n=516), 89 (n=486), vaccine (n=477), vaccines (n=477), 19, (n=362), disease (n=349). 
```
#create bigrams
#ANSWER: The top 10 bigrams were covid 19, public health, health care, 19 coronavirus, covid coronavirus, york times, language english, load date, publication type, and health departments. The next 10 were diseases disorders, infectious disease, rights reserved, disorders injuries, covid vaccine, video clip, usa 79, coronavirus vaccine, disease control, and disorders 90. 
```{r}
long_covid_bigrams <- articles_df |> 
  unnest_tokens(bigram, sentence, token = "ngrams", n = 2)

#Filter out stop words.

bigrams1 <- articles_df %>%
  unnest_tokens(bigram, sentence, token = "ngrams", n = 2) %>%
  filter(!is.na(bigram)) |> 
  separate(bigram, c("word1", "word2"), sep = " ")

bigrams2 <- bigrams1 %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word) |> 
  count(word1, word2, sort = TRUE)

head(bigrams2, 100)
```

#MEMO

#Summary of results: The top 10 words were covid with the highest count (n=1122), followed by health (n=952), 90 (n=730), 78 (n=573), people (n=516), 89 (n=486), vaccine (n=477), vaccines (n=477), 19, (n=362), disease (n=349). The top 10 bigrams were covid 19, public health, health care, 19 coronavirus, covid coronavirus, york times, language english, load date, publication type, and health departments. The next 10 were diseases disorders, infectious disease, rights reserved, disorders injuries, covid vaccine, video clip, usa 79, coronavirus vaccine, disease control, and disorders 90. 

#Immediately, I noticed that just looking at word counts was not helpful because there were many irrelevant words that showed up. Foror example, many numbers without context were included as frequent words. Even with words like "people" and "disease" that head high frequencies in the data, I realized that it is hard to interpret much from these results because there is no context for how those words were used in a sentence.
#The bigrams were helpful to see immediate context, but even then, I have a few more questions about how to interpret them. Many seemed nonsensical (i.e. "disorders 90" or "usa 79"). I wasn't sure what these meant, and was surprised to see as high of a frequency as I did for these bigrams in the data. Some bigrams were more expected as they are complete constructs that are constructed from 2 words (i.e. covid 19, public health, disease control, health departments and health care). The bigrams were significantly more helpful than looking at individual word counts to understand the specific construct that was being discussed in the articles.
#There are some irrelevant words in the bigrams which indicate that the bigrams included metadata (likely at the top or bottom of articles). For example, words like "publication type" or "load date." In future assignments, these should be cleaned up. 