---
title: "Metadata"
author: "Tennety"
date: "2026-09-15"
output: html_document
---

Class exercise to import your project metadata from the spreadsheet into R 

--Create a folder called "class_project" in your code folder 

--Save this file to the "class_project" folder 

--Move your Nexis spreadsheet to the folder 

--Create a project file on R Studio: File \| New Project \| \*\*\* select Existing folder \| "class_project" 

--After project.Rproj launches, then launch this file, metadata.Rmd, and complete the tutorial

**By doing this, you avoid the ring of hell where R can't find your data**

#load tidyverse, janitor, rio

```{r}

library(tidyverse)
library(janitor)
library(rio)
install.packages("readxl")
library(readxl)

```

#Import your data

```{r}

data <- read_excel("results_list_for_long COVID.XLSX")
View(data)

```

# Clean names

```{r}
data <- data |> 
  clean_names()

```

# check data types
```{r}
glimpse(data)
```


Next Steps: Count years, and create a ggplot

# Transform date into years. See tutorial
```{r}

data <- data |> 
  mutate(year = year(published_date))

```
```{r}
data |> 
  count(year)  
```
```{r}
data |> 
  count(publication_4) |> 
  arrange(desc(n))
```
```{r}
data |> 
  count(publication_location) |> 
  arrange(desc(n))
```
```{r}
data |> 
  count(publication_type_17) |> 
  arrange(desc(n))
```
```{r}
data_big <- data |> 
  mutate(length1 = as.numeric(length)) |> 
  filter (length1 >= 2000)
```


# Create a ggplot
```{r}
data |> 
  count(year) |> 
   ggplot(aes(x = year, y = n,
             fill = n)) +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
  scale_x_continuous(breaks=c(2015:2026)) + #fixes wonky x axis
  labs(title = "Years for Long COVID Reporting",
       subtitle = "Limited NexisUni Run",
       caption = "Graphic by Niveda Tennety, 9-16-2026",
       y="Number of Selected Articles",
       x="Year")



```

#Count publications

```{r}
data |> 
  count(publication_4) |> 
  arrange(desc(n))
```


#Visualize publications

```{r}
data |> 
  count(publication_4) |> 
   #ggplot(aes(y = publication_4, x = n,
             #fill = n)) +
  ggplot(aes(y = reorder (publication_4, n),x = n,
             fill = n)) +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
  labs(title = "Number of Articles by Publication",
       subtitle = "Limited NexisUni Run",
       caption = "Graphic by Niveda Tennety, 9-16-2026",
       y="Publication",
       x="Number of Selected Articles")
```


# NOTES ON FIXING COMPLEX Date COLUMN WITH HOURS AND MINUTES

```{r}
data_complex <- data |> 
  mutate(date = as.Date(strptime(published_date, format = "%B %d, %Y %A")))
```

#Next, I will examine average article length for different publications. 

#Reflection (code below): I found that CNN International and Fox News had the highest average length, but after looking at the table, I realized that this was not a useful statistic because the n value for each publication is so different. I have 33 articles from NYT but only 1 article or so from many sources, so this is not a valuable statistic. After looking at the table, and realizing that, I decided to not group by publication, and explore summaries overall of all the articles to get a sense of average length across all articles, for example. I thought this would be more valuable because rather than comparing publications in an unrigorous way (because of such stark differences in # of articles in each publication), I would be looking at the content as a whole. 

```{r}
glimpse(data)
data <- data |> 
  mutate(length1 = as.numeric(length))
```


```{r}
data %>% 
  select(publication_4, length1) %>% 
  group_by(publication_4) %>% 
  summarize(length1 = mean(length1, na.rm=TRUE)) |> 
  arrange(desc(length1))
```

#Summary statistics
#Reflection: The range of article length across this corpus is 280-6789 words. The median length is 1181 words, and the mean length is 1439 words. 
```{r}
summary(data$length1)
```
##I then looked up a framework on what constitutes short-form vs. long-form reporting. For my research project, I would vet the framework further, but just wanted to use one where I could practice R for the purpose of this class project. So, I chose the first framework that came up on a google search, which is the Nielsen Norman Group content strategy analysis. This search defines short-form articles as those with a word count of less than 1000 words, and long form articles as those with more than 1000 words. 

#First, I wanted to understand the average article length for publications.  
#Reflection: All publications except for 5 (Guardian, BI, USA Today, Axios, TOI) had average lengths above 1000. Because n values are low, this is also not a very meaningful statistic. 

```{r}
class(data$length)
data$length <-as.numeric(data$length)
```


```{r}
data %>% 
  select(publication_4, length) %>% 
 group_by(publication_4) %>% 
  summarize(Total=n(),
            Avg = mean(length, na.rm=TRUE)) %>% 
  arrange(desc(Avg))
```

#Then, for New York Times coverage specifically, because it had the largest n value, I wanted to understand how many articles were short-form vs. long-form based on this framework. 
#Answer: The new york times has 30 articles that are over 1000 words, and 3 articles that are less than 1000 words. 
```{r}
data |> 
  filter(publication_4 == "The New York Times") |> 
  filter (length>1000) |> 
  count()
```

